---
title: "Introduction to Quantitative Variables"
author: "P. Lombardo"
output:
  html_document: default
---

## Load packages and data sets
```{r}
library(tidyverse)
two_stores<-read.csv("data/two_stores.csv", header = T, 
                     stringsAsFactors = T)
source('source_files/general.R')
```


# Quantitative Variables

## Quantitative Summaries
Measures of center: `mean()`,  `median()`, and `mode()`.

Quick summaries
```{r}
#place code here
```

Making Summary Tables with `summarize()`
```{r}
two_stores %>%
    summarise(
    # list our summaries here: mean, median, sample size
    
    )
```

The summary table above combines both stores together into one large data set.  What if we want the same summaries, but separately for each store location? 

```{r}
# copy the summary table code above, and let's add to it.
```

**Exercise.** Look at the summary table above comparing the two stores. What can we say so far about the two stores? Do the sales in Beverly seem different than the sales in Boston?

*For those interested*: we can switch the rows and columns of these summary tables if we want, but it involves some trickery.  I wrote a small `flip_tibble()` function to accomplish this and it was "sourced" in the first code chunk with the `general.R` file.  Try it below!
```{r}
# copy and past the table, then add %>% flip_tibble() to the end

```


### Visualizing with Histograms
A simple histogram only has one aesthetic (`x`), and we use `geom_histogram()` as our geometry to produce the plot. Let's explore below by looking at a histogram of salmon weights.
```{r}
#Place code here to explore a basic histogram
ggplot(data = ...,
       aes(...))+
    geom_histogram()

# A little ugly at first, let's make some modifications...
    # bins = ... or 
    # breaks = seq(start, stop, by = step)
    # color = "white"
```


### Measures or Variation
While *measures of center* try to summarize where the "center" of the data values falls, *measures of variation* attempt to describe the degree to which the data values are spread out. Let's revisit our stores from the lens of *variation* in sales.

Here are the usual measures of variation: 

* standard deviation `sd()` (and its counterpart variance `var()`)
* inter-quartile range `IQR()` (*We will talk about IQR in the future lesson.*), and 
* range, defined as the maximum value minus the minimum. <br> There is no simple function for this, but also the least useful, in my opinion.  

**Exercise.** Copy and paste the summary table from above, but this time let's add in the range and std. deviation in our summary.
```{r}
# place the copy here and modify it.
```

So, ***what differentiates these two data sets?***

<answer>

Let's visualize this. Ideally, we want to compare histograms for each store (rather than all together like we did above).  We can do this in ggplot using a `facet_wrap(vars(...))` layer.
```{r}
ggplot(data = two_stores,
       aes(x = sales))+
    geom_histogram(bins = 10, color="white")
    # add to the plot to separate the histograms

# adding facet_grid(store~.) will also work, we may explore this another time.
```
**Challenge:** Can you make each store's histogram a different color?  Here's a hint... the fill of the histogram would depend on the data (specifically the `store` variable).


